### 2013-09-25 presentations-nccarf-2013/presentation-hanigan-final.tex

% Created 2013-06-26 Wed 11:16
\documentclass[bigger]{beamer}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{longtable}
\usepackage{float}
\mode<beamer>{\usetheme{Madrid}}
\usepackage{verbatim}
\usepackage{color}
\usepackage{amsmath,amsfonts,amssymb}
\providecommand{\alert}[1]{\textbf{#1}}

\title{Open Software - Restricted Data:  A Suicide/Climate Case Study.}
\author{Ivan Hanigan$^1$, David Fisher$^2$, Steven McEachern$^3$}
\date{June 26, 2013}
\hypersetup{
  pdfkeywords={},
  pdfsubject={},
  pdfcreator={Emacs Org-mode version 7.9.3f}}

\institute[NCEPH]{$^1$National Centre for Epidemiology and Population Health (ANU) \\ $^2$Information Technology Services (ANU) \\ $^3$Australian Data Archives (ANU)}
\begin{document}

\maketitle

% Org-mode is exporting headings to 3 levels.

\section{Aim}
\label{sec-1}
\begin{frame}
\frametitle{Aim}
\label{sec-1-1}

\begin{itemize}
\item Restrictions on data access have increased recently
\item Concerns regarding reproducibility of data analyses
\item Access to data and analytic software addresses the:\\
 \textbf{Replicability Crisis}, (Peng 2011, \emph{Science}, 334;6060)
\item We built a safe Sever/Client IT environment for this
\item We show a Case Study of Suicide and Climate Impacts research
\end{itemize}
\end{frame}
\section{Methods}
\label{sec-2}
\begin{frame}
\frametitle{Methods}
\label{sec-2-1}

\begin{figure}[!h]
\centering
\includegraphics[width=.65\textwidth]{opensoft.pdf}
\caption{1. System Design}
\label{fig:sys}
\end{figure}
\end{frame}
\section{Results}
\label{sec-3}
\begin{frame}
\frametitle{Results (Hanigan et al, 2012, \emph{PNAS}, 109;35)}
\label{sec-3-1}

\begin{footnotesize}
\begin{itemize}
\item {\color{red}Restricted Health and Climate data} and 
\item {\color{blue}Less Restricted Population data} 
\end{itemize}
(Colours refer to data storage and access rules shown in Figure 1).
\begin{eqnarray*}
        log({\color{red} O_{ijk}})  & = & s({\color{red}ExposureVariable})  + {\color{blue} OtherExplanators}  \\
        & &   + AgeGroup_{i} + Sex_{j} \\
        & &   + {\color{blue} SpatialZone_{k}}  \\
        & &  + sin(Time \times 2 \times \pi) + cos(Time \times 2 \times \pi) \\
        & &  + Trend \\
        & &   + offset({\color{blue} log(Pop_{ijk})})\\
\end{eqnarray*}
\end{footnotesize}
\begin{tiny}
\noindent Where:\\
        \indent ${\color{red}O_{ijk}}$ = Outcome (counts) by Age$_{i}$, Sex$_{j}$ and SpatialZone$_{k}$ \\
        \indent {\color{red}ExposureVariable} = Data with {\color{red}Restrictive Intellectual Property~(IP)} \\
        \indent {\color{blue}OtherExplanators} = Other {\color{blue}Less Restricted}  Explanatory variables \\
        \indent s( ) = penalized regression splines \\
        \indent ${\color{blue} SpatialZone_{k}}$  = {\color{blue} Less Restricted} data representing the $SpatialZone_{k}$  \\
        \indent Trend = Longterm smooth trend(s) \\
        \indent ${\color{blue}Pop_{ijk}}$ = interpolated Census populations, by time in each group\\
\end{tiny}
\end{frame}
\begin{frame}
\frametitle{Future (Bambrick et al, 2008, Garnaut Review)}
\label{sec-3-2}

\begin{footnotesize}
$$Y_{ijk}=\sum_{lm}(e^{(\beta_{ijk} \times {\color{red} X_{lm}})} - 1) \times {\color{red}BaselineRate_{jkl}} \times {\color{blue} Population_{jklm}}$$
\noindent Where:\\
$\beta_{ijk}$ = the ExposureVariable coefficient for zone$_i$, age$_j$ and sex$_{k}$ \\
${\color{red}X_{lm}}$ = Projected Future ExposureVariables {\color{red} with Restrictive IP} \\
{\color{red}BaselineRate$_{jkl}$} = {\color{red}avgDeathsPerTime}/{\color{blue}avgPopPerTime} in age$_j$, sex$_k$ and zone$_l$ \\
{\color{blue}Population$_{jklm}$} = projected populations by age$_j$, sex$_k$, zone$_l$ and time$_m$ {\color{blue} (With Less Restrictions)}\\

\end{footnotesize}
\end{frame}
\section{Conclusion}
\label{sec-4}
\begin{frame}
\frametitle{Conclusion}
\label{sec-4-1}

This system:
\begin{itemize}
\item Enables data analysis in a safe environment
\item Allows comparison of multiple climate scenarios and assumptions
\item Demonstrated with a Climate/Health Impact Assessment
\end{itemize}
\begin{itemize}
\begin{large}
\item And this is Reproducible
\end{large}
\end{itemize}
\end{frame}
\section{Acknowledgements}
\label{sec-5}
\begin{frame}
\frametitle{Acknowledgements}
\label{sec-5-1}

\includegraphics[width=4cm]{ANU_LOGO_cmyk_56mm.png}
\includegraphics[width=3cm]{andslogo.pdf}
\includegraphics[width=3cm]{deptlogo.pdf} \\
\begin{footnotesize}
This project is supported by the Australian National Data Service through the National Collaborative Research Infrastructure Strategy Program and the Education Investment Fund (EIF) Super Science Initiative.

More information from:
\begin{itemize}
\item \texttt{ivan.hanigan@gmail.com}
\item \texttt{http://opensoftware-restricteddata.github.io}
\end{itemize}
\end{footnotesize}
\end{frame}
\section{References}
\label{sec-6}
\begin{frame}
\frametitle{References}
\label{sec-6-1}

\begin{footnotesize}
\begin{thebibliography}{1}

\bibitem{Peng2011}
Roger~D Peng.
\newblock {Reproducible research in computational science.}
\newblock {\em Science (New York, N.Y.)}, 334(6060):1226--7, December 2011.

\bibitem{Hanigan2012b}
I.~C. Hanigan, C.~D. Butler, P.~N. Kokic, and M.~F. Hutchinson.
\newblock {Suicide and drought in New South Wales, Australia, 1970-2007}.
\newblock {\em Proceedings of the National Academy of Sciences}, pages
  1112965109--, August 2012.

\bibitem{Climate2008}
Hilary~J Bambrick, Keith B~G Dear, RE~Woodruff, Ivan~Charles Hanigan, and
  Anthony~J McMichael.
\newblock {The impacts of climate change on three health outcomes:
  temperature-related mortality and hospitalisations, salmonellosis and other
  bacterial gastroenteritis, and population at risk from dengue.}
\newblock Technical report, Garnaut Climate Change Review, Canberra, 2008.

\end{thebibliography}
\end{footnotesize}
\end{frame}

\end{document}

### 2013-10-04 presentations-india-mssrf/presentation-mssrf.tex
half done

--- 2013-09-25 presentations-nccarf-2013/presentation-hanigan-final.tex
@@
-% Created 2013-06-26 Wed 11:16
+% Created 2013-10-04 Fri 18:48
 \documentclass[bigger]{beamer}
@@
 
-\title{Open Software - Restricted Data:  A Suicide/Climate Case Study.}
+\title{Climate Change Health Impact Assessments:  Farmer Suicide and Drought Case Study.}
 \author{Ivan Hanigan$^1$, David Fisher$^2$, Steven McEachern$^3$}
-\date{June 26, 2013}
+\date{\today}
 \hypersetup{
@@
 \begin{itemize}
-\item Restrictions on data access have increased recently
-\item Concerns regarding reproducibility of data analyses
-\item Access to data and analytic software addresses the:\\
- \textbf{Replicability Crisis}, (Peng 2011, \emph{Science}, 334;6060)
-\item We built a safe Sever/Client IT environment for this
-\item We show a Case Study of Suicide and Climate Impacts research
+\item General tools for Climate Change Health Impact Assessments (CCHIA)
+\item Enhanced capacity for experimentation, reviews, revisions and re-iterations
+\end{itemize}
+
+Current approach:
+\begin{itemize}
+\item historical baseline exposure-response functions, control for some covariates
+\item use response function with changed exposures and population at risk
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Methods}
+\frametitle{Historical (Hanigan et al, 2012, \emph{PNAS}, 109)}
 \label{sec-2-1}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=.65\textwidth]{opensoft.pdf}
-\caption{1. System Design}
-\label{fig:sys}
-\end{figure}
-\end{frame}
-\section{Results}
-\label{sec-3}
-\begin{frame}
-\frametitle{Results (Hanigan et al, 2012, \emph{PNAS}, 109;35)}
-\label{sec-3-1}
 
@@
 \begin{itemize}
-\item {\color{red}Restricted Health and Climate data} and 
+\item {\color{red}Restricted Health and Drought data} and 
 \item {\color{blue}Less Restricted Population data} 
@@
 \begin{frame}
+\frametitle{Historical (Hanigan et al, 2012, \emph{PNAS}, 109)}
+\label{sec-2-2}
+
+\begin{itemize}
+\item 38 years suicide rates with drought by 11 regions, age and sex
+\item Estimated 9\% in rural males aged 30-49 due to drought over the period
+\item Increased for rural males 10-29 y
+\item Association with hot temp + spring
+\end{itemize}
+\end{frame}
+\begin{frame}
 \frametitle{Future (Bambrick et al, 2008, Garnaut Review)}
-\label{sec-3-2}
+\label{sec-2-3}
 
@@
 \end{frame}
+\section{Results}
+\label{sec-3}
+\begin{frame}
+\frametitle{Drought-suicide response function}
+\label{sec-3-1}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=.5\textwidth]{Figure1.png}
+\caption{Figure1.png}
+\label{fig:Figure1.png}
+\end{figure}
+\end{frame}
+\begin{frame}
+\frametitle{Criticism}
+\label{sec-3-2}
+
+This model is too static, reductionist, reality is more complex. Need to work more on interactions with non-climate factors especially: 
+\begin{itemize}
+\item Natural capital
+\item Financial capital
+\item Social capital
+\item Physical capital and
+\item Human capital
+\end{itemize}
+\end{frame}
 \section{Conclusion}
@@
 
-This system:
 \begin{itemize}
-\item Enables data analysis in a safe environment
-\item Allows comparison of multiple climate scenarios and assumptions
-\item Demonstrated with a Climate/Health Impact Assessment
-\end{itemize}
-\begin{itemize}
-\begin{large}
-\item And this is Reproducible
-\end{large}
+\item Drought is related to increased suicide risk in Australia
+\item Future Drought associated deaths can be calculated
+\item These estimates will be very uncertain, contentious and difficult to justify
+\item New technology is needed to enable rigorous and transparent exploration
 \end{itemize}
@@
 \includegraphics[width=4cm]{ANU_LOGO_cmyk_56mm.png}
-\includegraphics[width=3cm]{andslogo.pdf}
+\includegraphics[width=2cm]{andslogo.pdf}
 \includegraphics[width=3cm]{deptlogo.pdf} \\
@@
 
-More information from:
-\begin{itemize}
-\item \texttt{ivan.hanigan@gmail.com}
-\item \texttt{http://opensoftware-restricteddata.github.io}
-\end{itemize}
+More information from \texttt{ivan.hanigan@gmail.com} or at \texttt{[[http://opensoftware-restricteddata.github.io][http://opensoftware-restricteddata.github.io]]}
+
 \end{footnotesize}

### 2013-10-04 presentations-india-mssrf/presentation-mssrf.tex
add mssrf cc pres

@@
-% Created 2013-10-04 Fri 18:48
+% Created 2013-10-05 Sat 01:18
 \documentclass[bigger]{beamer}
@@
 \usepackage{amsmath,amsfonts,amssymb}
+\AtBeginSection[]{\begin{frame}<beamer>\frametitle{Topic}\tableofcontents[currentsection]\end{frame}}
 \providecommand{\alert}[1]{\textbf{#1}}
@@
 
+Climate Change Health Impact Assessments (CCHIA) need
 \begin{itemize}
-\item General tools for Climate Change Health Impact Assessments (CCHIA)
-\item Enhanced capacity for experimentation, reviews, revisions and re-iterations
+\item Workflow tools for data management and analysis
+\item Enhanced capacity for experimentation, reviews, revisions
 \end{itemize}
 
-Current approach:
+General approach:
 \begin{itemize}
-\item historical baseline exposure-response functions, control for some covariates
-\item use response function with changed exposures and population at risk
+\item historical exposure-response functions, control covariates
+\item future response with changed exposures and population at risk
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{Motivating Case Study: Climate/Suicide}
+\label{sec-1-2}
+
+\begin{itemize}
+\item Suicide has been linked to climate in a variety of studies
+\item Climate change impact on mental health is a gap in knowledge
+\item Use methods to analyse the relationship between Drought/Suicide
+\item Estimate future Climate Change impacts
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{New Tools are Needed}
+\label{sec-1-3}
+
+\begin{itemize}
+\item Restrictions on access to suicide data have increased recently.
+\item Growing concern about the \textbf{Replicability Crisis}, (Peng 2011, \emph{Science}, 334).
+\item Access to data and analytic software addresses this.
+\item We built a safe Sever/Client IT environment.
 \end{itemize}
@@
 \begin{frame}
-\frametitle{Historical (Hanigan et al, 2012, \emph{PNAS}, 109)}
+\frametitle{System Design}
 \label{sec-2-1}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=.65\textwidth]{opensoft.pdf}
+\caption{1. System Design}
+\label{fig:sys}
+\end{figure}
+\end{frame}
+\begin{frame}
+\frametitle{Linux Cluster}
+\label{sec-2-2}
+
+\begin{itemize}
+\item National Research Cloud (www.nectar.org.au/research-cloud)
+\item Centos 6.4 (www.centos.org)
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{PostGIS Database (The Brawn)}
+\label{sec-2-3}
+
+\begin{itemize}
+\item PostgreSQL 9.2 (www.postgresql.org)
+\item PostGIS 2.0 (postgis.refractions.net)
+\end{itemize} 
+\end{frame}
+\begin{frame}
+\frametitle{Analysis (The Brains)}
+\label{sec-2-4}
+
+\begin{itemize}
+\item R language for statistical computing (www.r-project.org)
+\item Rstudio server (www.rstudio.com/)
+\item OpenGeo Suite (opengeo.org)
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{Information Management}
+\label{sec-2-5}
+
+\begin{itemize}
+\item Projects,UsersDB Oracle XE APEX (www.oracle.com)
+\item Data Catalogue (assda.anu.edu.au/ddiindex.html)
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{The Client Side}
+\label{sec-2-6}
+
+\begin{itemize}
+\item The Kepler Project (www.kepler-project.org)
+\item pgAdmin (www.pgadmin.org)
+\item Git Version Control and GitHub (github.com)
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{Case Study: Historical}
+\label{sec-2-7}
 
@@
 \begin{frame}
-\frametitle{Historical (Hanigan et al, 2012, \emph{PNAS}, 109)}
-\label{sec-2-2}
+\frametitle{Case Study: Historical}
+\label{sec-2-8}
 
+Hanigan et al, 2012, \emph{PNAS}, 109:
 \begin{itemize}
@@
 \begin{frame}
-\frametitle{Future (Bambrick et al, 2008, Garnaut Review)}
-\label{sec-2-3}
+\frametitle{Case Study: Future}
+\label{sec-2-9}
 
+Bambrick et al, 2008, Garnaut Review:
 \begin{footnotesize}
@@
 \centering
-\includegraphics[width=.5\textwidth]{Figure1.png}
+\includegraphics[width=.45\textwidth]{Figure1.png}
 \caption{Figure1.png}
@@
 \end{frame}
+\section{Discussion}
+\label{sec-4}
 \begin{frame}
 \frametitle{Criticism}
-\label{sec-3-2}
+\label{sec-4-1}
 
@@
 \section{Conclusion}
-\label{sec-4}
+\label{sec-5}
 \begin{frame}
 \frametitle{Conclusion}
-\label{sec-4-1}
+\label{sec-5-1}
 
@@
 \item These estimates will be very uncertain, contentious and difficult to justify
-\item New technology is needed to enable rigorous and transparent exploration
+\item Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration
+\end{itemize}
+\end{frame}
+\begin{frame}
+\frametitle{Conclusion}
+\label{sec-5-2}
+
+This system:
+\begin{itemize}
+\item Enables data analysis in a safe environment
+\item Allows comparison of multiple climate scenarios and assumptions
+\item Demonstrated with a Climate/Health Impact Assessment
+\item This is Reproducible
 \end{itemize}
@@
 \section{Acknowledgements}
-\label{sec-5}
+\label{sec-6}
 \begin{frame}
 \frametitle{Acknowledgements}
-\label{sec-5-1}
+\label{sec-6-1}
 
@@
 \section{References}
-\label{sec-6}
+\label{sec-7}
 \begin{frame}
 \frametitle{References}
-\label{sec-6-1}
+\label{sec-7-1}
 

### 2013-10-04 presentations-nccarf-2013/poster.tex
add mssrf cc pres

\documentclass[final]{beamer}
\usetheme{RJH}
\usepackage[orientation=portrait,size=a1,scale=1.4,debug]{beamerposter}
\usepackage[absolute,overlay]{textpos}
\usepackage{amsmath,amsfonts,amssymb}
\usepackage{color}
\setlength{\TPHorizModule}{1cm}
\setlength{\TPVertModule}{1cm}

%\includegraphics[height=2cm,width=3cm]{ANU_LOGO_cmyk_56mm.pdf}
\title{Open Software - Restricted Data: The Suicide/Climate Case}
\author{Ivan Hanigan$^1$, David Fisher$^2$, Steven McEachern$^3$}
\footer{More information from \texttt{ivan.hanigan@gmail.com} or at \texttt{http://opensoftware-restricteddata.github.io}}
\date{}

\begin{document}

%\begin{textblock}{6}(0.,0.2)
%\includegraphics[width=3cm]{ANU_LOGO_cmyk_56mm.png}
%\end{textblock}

\begin{frame}{}

\begin{textblock}{17}(2,5)
\begin{block}{Restricted Data is Safe}
\begin{scriptsize}
Restrictions on access to confidential health data have increased recently. Enabling safe access to data and analytic software is needed to address the {\color{red}\textbf{Replicability Crisis}} (Peng 2011).  We present an environment for analysing restricted data using open software. The system is described in Figure~\ref{fig:sys} and a Case Study of the historical association of suicides with climate (and extrapolation of this under climate change/adaptation scenarios) in the bottom half of the poster. These tools allow users to access restricted data; protects confidentiality and allows use of open software for reproducibility (King~1995).  
\end{scriptsize}
\end{block}

\begin{block}{Restrictive IT Environments}
\begin{scriptsize}
Previous solutions to this challenge make access so restricted that usability is compromised. We aimed to build a collection of tools for the conduct of many types of health and social science research. The starting point for users is the data catalogue, which provides for finding data available from the store of {\color{red} Restricted data} and {\color{blue}Less~Restricted data}  for approved use. Once data are discovered, the researcher has capacity to manipulate the datasets on the secure server. The PostgreSQL database integrates and Geoserver visualises, while statistical tools are available in the R-studio server browser.
\end{scriptsize}

\end{block}
\end{textblock}

\begin{textblock}{17}(21.25,5)
\begin{block}{Server-Client Architecture}
\begin{figure}[!h]
\centering
\includegraphics[width=\textwidth]{opensoft.pdf}
\caption{1. System Design}
\label{fig:sys}
\end{figure}
\end{block}

\begin{block}{The Stack}
\begin{tiny}
\textbf{Hardware}:
\begin{itemize}
\item National Research Cloud \emph{http://www.nectar.org.au/research-cloud/}
\item Centos 6.4 \emph{www.centos.org/}
\end{itemize}

\textbf{Database (The Brawn)}:
\begin{itemize}
\item PostgreSQL 9.2 \emph{http://www.postgresql.org/}
\item PostGIS 2.0 \emph{http://postgis.refractions.net/}
\end{itemize}

\textbf{Analysis (The Brains)}:
\begin{itemize}
\item R language for statistical computing \emph{http://www.r-project.org/}
\item Rstudio server \emph{www.rstudio.com/‎}
\item OpenGeo Suite \emph{http://opengeo.org/}
\end{itemize}

\textbf{Information Management:}
\begin{itemize}
\item Projects,UsersDB Oracle XE APEX \emph{www.oracle.com}
\item Data Catalogue \emph{http://assda.anu.edu.au/ddiindex.html}
\end{itemize}

\textbf{The Client Side}:
\begin{itemize}
\item The Kepler Project \emph{https://kepler-project.org/‎}
\item pgAdmin \emph{www.pgadmin.org/‎}
\item Git Version Control and GitHub \emph{https://github.com/}
\end{itemize}

\end{tiny}
\end{block}

\end{textblock}

\begin{textblock}{17}(40.5,5)

\begin{block}{Reproducibility}
\begin{scriptsize}
Such analytical tools will enhance the ability of adaptive management practitioners to assess the potential influence of adaptations. The use of the system shows the ease with which multiple data sources (some restricted) can be analysed in a secure way using open software. This will build capacity to answer complex research questions and compare multiple climate change scenarios or adaptation assumptions; achieving simultaneous vision of potential future outcomes from different standpoints.
\end{scriptsize}

\end{block}

\begin{block}{How it works}
\begin{scriptsize}
\begin{itemize}
\item Open a web browser / log on to the catalogue / find data
\item In the web browser log on to Rstudio server
\item Connect to database / query datasets / join / subset / transform
\item Get data to your Rstudio server workspace and analyse
\item Commit all code to GitHub, download resulting dataset and reports 
\end{itemize}
\end{scriptsize}
\end{block}

\begin{block}{Summary}
\begin{scriptsize}
This system:
\begin{itemize}
\item Assists rigorous data management practices
\item Enables multiple data sources (some restricted) to be analysed
\item Storage of data is secure
\item Analytic code is made available as open software
\end{itemize}

\begin{itemize}
\begin{large}
\item Enhances reproducibility
\end{large}
\end{itemize}
\end{scriptsize}
\end{block}

\end{textblock}

\begin{textblock}{55.5}(2,45)

\begin{block}{A Case Study}
\end{block}

\end{textblock}

\begin{textblock}{17}(2,47)

\begin{block}{Exposure/Response function}
\begin{tiny}

Historical association between Suicide and Climate Variables were established in a Poisson time-series model (Hanigan et al 2012) using:
\begin{itemize}
\item {\color{red}Restricted Health and Drought data} and 
\item {\color{blue}Less Restricted Population data} 
\end{itemize}

(Colours refer to data storage and access rules shown in Figure ~\ref{fig:sys}).

\begin{eqnarray*}
        log({\color{red} O_{ijk}})  & = & s({\color{red}ExposureVariable})  + {\color{blue} OtherExplanators}  \\
        & &   + AgeGroup_{i} + Sex_{j} \\
        & &   + {\color{blue} SpatialZone_{k}}  \\
        & &  + sin(Time \times 2 \times \pi) + cos(Time \times 2 \times \pi) \\
        & &  + Trend \\
        & &   + offset({\color{blue} log(Pop_{ijk})})\\
        \end{eqnarray*}
        \noindent Where:\\
        \indent ${\color{red}O_{ijk}}$ = Outcome (counts) by Age$_{i}$, Sex$_{j}$ and SpatialZone$_{k}$ \\
        \indent {\color{red}ExposureVariable} = Data with {\color{red}Restrictive Intellectual Property~(IP)} \\
        \indent {\color{blue}OtherExplanators} = Other {\color{blue}Less Restricted}  Explanatory variables \\
        \indent s( ) = penalized regression splines \\
        \indent ${\color{blue} SpatialZone_{k}}$  = {\color{blue} Less Restricted} data representing the $SpatialZone_{k}$  \\
        \indent Trend = Longterm smooth trend(s) \\
        \indent ${\color{blue}Pop_{ijk}}$ = interpolated Census populations, by time in each group\\
\end{tiny}

\end{block}

\begin{block}{Climate Change Scenarios}
\begin{tiny}
We can use methods like Bambrick et al 2008 to estimate Climate Change Health Impacts:

$$Y_{ijk}=\sum_{lm}(e^{(\beta_{ijk} \times {\color{red} X_{lm}})} - 1) \times {\color{red}BaselineRate_{jkl}} \times {\color{blue} Population_{jklm}}$$
Where:\\
$\beta_{ijk}$ = the ExposureVariable coefficient for zone$_i$, age$_j$ and sex$_{k}$ \\
${\color{red}X_{lm}}$ = Projected Future ExposureVariables {\color{red} with Restrictive IP} \\
{\color{red}BaselineRate$_{jkl}$} = {\color{red}avgDeathsPerTime}/{\color{blue}avgPopPerTime} in age$_j$, sex$_k$ and zone$_l$ \\
{\color{blue}Population$_{jklm}$} = projected populations by age$_j$, sex$_k$, zone$_l$ and time$_m$ {\color{blue} (With Less Restrictions)}\\

\end{tiny}
\end{block}
\end{textblock}
\begin{textblock}{17}(21.25,47)

\begin{block}{Suicide and Temperature}
\begin{tiny}
The association of suicide and maximum temperature anomalies is shown in Figure ~\ref{fig:suigam}.  This can be used to estimate future climate impacts using future climate scenarios {\color{red} (Data with Restrictive IP)}  and population at risk {\color{blue} (Less Restricted data)}.
\end{tiny}
\begin{figure}[!h]
\centering
\includegraphics[width=.6\textwidth]{tmaxanom2.pdf}
\caption{2. Suicide/Temperature}
\label{fig:suigam}
\end{figure}
\end{block}

\begin{block}{Conclusions}
\begin{scriptsize}
This system:
\begin{itemize}
\item Enables data analysis in a safe environment
\item Allows comparison of multiple climate scenarios and assumptions
\item Demonstrated with a Climate/Health Impact Assessment
\end{itemize}
\begin{itemize}
\begin{large}
\item And this is Reproducible
\end{large}
\end{itemize}
\end{scriptsize}
\end{block}
\end{textblock}

\begin{textblock}{17}(40.5,47)

\begin{block}{References}
\begin{tiny}
\begin{thebibliography}{1}

\bibitem{Peng2011}
Roger~D. Peng.
\newblock {Reproducible research in computational science.}
\newblock {\em Science}, 334(6060), December 2011.

\bibitem{King1995}
Gary King.
\newblock {Replication, replication}.
\newblock {\em Political Science and Politics}, 28(3), September 1995.

\bibitem{Hanigan2012b}
Ivan~C. Hanigan, Colin~D. Butler, Phillip~N. Kokic, and Michael~F.~Hutchinson.
\newblock {Suicide and drought in New South Wales, Australia, 1970-2007}.
\newblock {\em Proceedings of the National Academy of Sciences}, 109(35), August~2012.

\bibitem{Climate2008}
Hilary~J. Bambrick, Keith~B.G. Dear, Rosalie~E. Woodruff, Ivan~C.~Hanigan, and
  Anthony~J. McMichael.
\newblock {The impacts of climate change on three health outcomes:
  temperature-related mortality and hospitalisations, salmonellosis and other
  bacterial gastroenteritis, and population at risk from dengue.}
\newblock Technical report, Garnaut Climate Change Review, Canberra, 2008.

\end{thebibliography}
\end{tiny}
\end{block}

\begin{block}{Acknowledgements}    
\begin{tiny}
\begin{itemize}
\item $^1$National Centre for Epidemiology and Population Health, ANU
\item $^2$Information Technology Services, ANU
\item $^3$Australian Data Archive, ANU 
\end{itemize}
\includegraphics[width=7cm]{ANU_LOGO_cmyk_56mm.png}
\includegraphics[width=4cm]{andslogo.pdf}
\includegraphics[width=6cm]{deptlogo.pdf} \\
This project is supported by the Australian National Data Service through the National Collaborative Research Infrastructure Strategy Program and the Education Investment Fund (EIF) Super Science Initiative.
\end{tiny}
\end{block}

\end{textblock}

\end{frame}
\end{document}

### 2015-11-03 report1_high_level/manuscript.Rmd
add report manuscript

---
title: 'Open Software - Restricted Data:  A Suicide/Climate Case Study.'
author:  
- name: Ivan C. Hanigan
  affilnum: 1
  email: ivan.hanigan@anu.edu.au  
- name: David Fisher
  affilnum: 2
- name: Steven McEachern
  affilnum: 3
affiliation:
- affilnum: 1
  affil: National Centre for Epidemiology and Population Health (ANU) 
- affilnum: 2
  affil: Information Technology Services (ANU)
- affilnum: 3
  affil: Australian Data Archives (ANU)
header-includes:
  - \usepackage{graphicx}
  - \usepackage{url}   
output:
  pdf_document:
    fig_caption: yes
    keep_tex: yes
    number_sections: yes
    template: components/manuscript.latex
  html_document: null
  word_document: null
fontsize: 11pt
capsize: normalsize
csl: meemodified.csl
documentclass: article
classoption: a4paper
spacing: singlespacing
linenumbers: no
bibliography: references.bib
abstract: no
---
<!--
%\VignetteEngine{knitr::rmarkdown}
%\VignetteIndexEntry{Example Manuscript}
-->
```{r, eval = F, echo = F}
library(rmarkdown)
library(knitr)
library(knitcitations)
library(bibtex)
cleanbib()
#rm("bib")
#options("cite_format"="pandoc")
cite_options(citation_format = "pandoc", check.entries=FALSE)

dir()
render("manuscript.Rmd", "pdf_document")
#browseURL("manuscript.pdf")
```
```{r, echo = F, results = 'hide'}
# load
if(!exists("bib")){
bib <- read.bibtex("~/references/library.bib")

for(bibkey in c("SarathiBiswas2012",
  "Mcmichael2002a", "Gelman2013"
)){
bib[[bibkey]]$url <- gsub("\\{\\\\_\\}","_", bib[[bibkey]]$url)
bib[[bibkey]]$url <- gsub("\\{~\\}","~", bib[[bibkey]]$url)
}

}
```
```{r setup, include=FALSE, echo=FALSE}
#Put whatever you normally put in a setup chunk.
#I usually at least include:
#devtools::install_github("manuscriptPackage","jhollist")
#library("manuscriptPackage")
#Didn't do that here to expedite building of the example vignette
library("knitr")

opts_chunk$set(dev = 'pdf', fig.width=6, fig.height=5)

# Table Captions from @DeanK on http://stackoverflow.com/questions/15258233/using-table-caption-on-r-markdown-file-using-knitr-to-use-in-pandoc-to-convert-t
#Figure captions are handled by LaTeX

knit_hooks$set(tab.cap = function(before, options, envir) {
                  if(!before) { 
                    paste('\n\n:', options$tab.cap, sep='') 
                  }
                })
default_output_hook = knit_hooks$get("output")
knit_hooks$set(output = function(x, options) {
  if (is.null(options$tab.cap) == FALSE) {
    x
  } else
    default_output_hook(x,options)
})
```

```{r analysis , include=FALSE, echo=FALSE, cache=FALSE}
#All analysis in here, that way all bits of the paper have access to the final objects
#Place tables and figures and numerical results where they need to go.
```

<!-- Abstract is being wrapped in latex here so that all analysis can be run in the chunk above and the results reproducibly referenced in the abstract. -->

\singlespace

\vspace{2mm}\hrule
\paragraph*{Background:} This essay was written to accompany the material presented as a
speedtalk and poster at the National Climate Change Adaptation
Research Facility Conference 'Climate Adaptation knowledge and
partnership', June 2013, Sydney.  The poster and slideshow are both available to download from this website: [http://opensoftware-restricteddata.github.io/presentations-nccarf-2013/](http://opensoftware-restricteddata.github.io/presentations-nccarf-2013/)

\paragraph*{Methods:} The paper reports on a project to build tools and procedures for enhancing open and transparent analysis of restricted datasets.
Some datasets such as suicide or climate
change scenarios need to be accessed in a restricted way.  On the
other hand scientists need to make their methods, models and
assumptions transparent and available for scientific debate even
though the datasets may require authorisation to access.

\paragraph*{Results:} We built a safe Server/Client Computational
Environment for using open software with restricted data.  We
demonstrate the use of this system using drought and suicide as a case study.  
We describe the potential use of this system in modelling climate change scenarios.

\paragraph*{Conclusions:} The project shows that restricted data and
open software can be used in an appropriate way to further the
progress of scientific enquiry.

\vspace{3mm}\hrule
\doublespace

# Background 
## Open software for restricted data

Some datasets such as sensitive personal information about suicide or
climate change scenarios with protected intellectual property need to
be accessed in a restricted way.  In the context of Reproducible
Research (RR) methods, models and assumptions need to be made
transparent and available for scientific debate even though the
datasets may require authorisation to access `r citep(bib[["Peng2011"]])`.

Restrictions around access to data have increased recently in
Australia, especially to the national mortality database after the
discovery of an incident in which researchers at the University of
Queensland School of Population Health, dissatisfied with the time
it was taking the data custodians to provide data, had hacked into
information about deaths in Australia `r citep(bib[["OKeefe2007"]])`.  
The subsequent investigations lead to
a wide ranging modification to the procedures for approval and
provision of these data that make the access much more restricted.

The Replicability Crisis `r citep(bib[["Peng2011"]])` has emerged, reducing public confidence in statements by scientists. The true extent of the problem may turn out to be overstated `r citep(bib[["Jager2014"]])` however the concern should be addressed to avoid the problems a lack of confidence in scientific publications would entail. Appropriate access to data and analytic software addresses this issue.   We investigated available workflow tools for data management and analysis and implemented a range of these products on our server.  This server has enhanced our capacity for experimentation, reviews, revisions and extensions of work in this field.  We present the results of this project and report that it has streamlined access to population health and environmental data for analysis.

## Motivating case study: Climate/suicide

Suicide has been linked to climate in a variety of studies.
The climate change impact on mental health is a gap in knowledge.
The motivating case for this project was to analyse the relationship between Drought and Suicide. 
The historical exposure-response function can be used to estimate future climate change impacts.

There has been substantial public interest within Australia in recent decades of the putative relationship between drought and rural mental health, including suicide. The topic has frequently been raised by the media, by rural politicians and by mental health support groups `r citep(bib[["AustralianB2006"]])`. There have also been media reports in India indicating substantial concerns about drought and rural suicide in that country, including `r citep(bib[["SarathiBiswas2012"]])` from May 2012.

The number of studies that have examined the relationship between suicide and drought is limited. However, many papers explore links between suicide and climate variables other than drought (such as temperature) and there are two major reviews papers available of the literature on climatic influences on suicide `r citep(c(bib[["Dixon2009"]],bib[["Deisenhammer2003b"]]))`. However, very few studies have investigated the "dryer than average conditions" that is drought specifically.

There are several mechanisms through which unusually low rainfall, especially if exacerbated by increased soil dryness due to higher temperatures may increase the suicide rate. First, droughts increase the financial stress on farmers and farming communities (even if partially compensated by drought relief welfare payments). Such difficulty may occur in conjunction with other economic stresses, such as rising interest rates, falling commodity prices, or an unfavourable foreign exchange rate.

As mentioned the number of studies that have examined the relationship
between suicide and drought is limited to only a handful `r citep(c(bib[["Nicholls2006c"]], bib[["Page2002c"]], bib[["Guiney2012a"]], bib[["Hanigan2012e"]]))` A  systematic
literature review of the Health Effects of Drought found little other
evidence for the putative causal effect of drought on suicide `r citep(bib[["Stanke2013"]])`.

# Methodology

The approach we took to meet the challenge of analysing restricted suicide and climate change scenario data in a safe environment was to
build a new hardware and software stack to using Open-Source software.  We based our
planning on the realisation that there is a growing need of these
technologies in the context of Reproducible Research (RR).  This
requires that methods, models and assumptions need to be made
transparent and available for scientific debate even though the
datasets may require authorisation to access.  This is true not just
in Health data, but also including the context 
of data with restrictive Intellectual Property and licence requirements.

### Open software for restricted data

To develop an over-all view of
the issue and analyse the dimensions of the problem we spent the
initial phase of the project conceptualising an overall "Rich Picture"
of the issue, and focused on identifying risks that the project might
face. Several papers that describe similar systems were reviewed `r citep(c(bib[["Evans2012"]], bib[["Fleming2014"]]))` and 
several recommendations from these papers were adopted in our system.

Our design responds directly to the primary threat of unintentional release of
sensitive data so we decided to build a safe Server/Client environment
for analysts to develop their software in an open way, while ensuring
the safety of the datasets.  Other risks we identified were in
relation to the provision of the server hardware and we were able to
take advantage of the Nectar Research Cloud for virtualised services.

Then we defined the scope and quality of the project outcomes that we
were aiming to deliver.  The fact that restrictions around access to
data have increased recently, coupled with arguments that appropriate
access to analytic software is needed to address the Replicability
Crisis meant that the scope of this project was very broad.  we also
explored the ambitions of our analysts to support their publishable
outputs with open software.  Given that examples of un-replicable work
has spread even to the results published in top journals such as
Nature and Science, the scope we decided to set for this project was
for very high levels of open-ness for the evidence being presented for
peer-review along with very high levels of restriction on access to
the data. Luckily however we were able to rule out the need for the
extreme level of restriction such as getting Defence Science and
Technology Organisation (DSTO) accreditation for the security of these
servers against malicious hacking attacks.

We also looked at the workflow system Kepler to assess it's utility
for providing access to the data, but found that there were a lot of
limitations at the time `r citep(bib[["Curcin2008"]])` and decided
that the R environment for statistical computing and graphics would be
the platform we would focus on.

During the next phase of the project we dealt with issues of the costs
associated with developing the software and hosting the hardware at
different locations, as well as the time needed to test and get user
acceptance on the services.  Throughout the project we have had to deal
with lack of resources, especially in terms of Software
Engineering skills for the Client-side user interface and Linux
Systems Administration support for the Server-side backend.

The results of the IT Infrastructure part of this project are described next.

## System requirements

In this case study we utilise Virtual Machines (VMs) in the Cloud. Our
system requires two VMs so that the storage and processing of data can
be compartmentalised, with various benefits.  A high level overview of
the system is shown in Figure \ref{fig:sys}.  Full details including
Linux commands and configuration specifications are available online at
[http://opensoftware-restricteddata.github.com](http://opensoftware-restricteddata.github.io).

\begin{figure}[!h]
\centering
\includegraphics[width=.85\textwidth]{opensoft.pdf}
\caption{High Level Schematic System Design, colours indicate restrictions (red), open (blue)}
\label{fig:sys}
\end{figure}

## Software selection process
We researched a variety of systems and found the following set-up worked best for us.

### Linux cluster

- National Research Cloud [www.nectar.org.au/research-cloud](www.nectar.org.au/research-cloud)
- Centos 6.4 [www.centos.org](www.centos.org)

### PostGIS database

- PostgreSQL 9.2 [www.postgresql.org](www.postgresql.org)
- PostGIS 2.0 [http://postgis.refractions.net](http://postgis.refractions.net)

### Analysis

-  R language for statistical computing [www.r-project.org](www.r-project.org)
-  Rstudio server [www.rstudio.com](www.rstudio.com)
-  OpenGeo Suite [http://opengeo.org](http://opengeo.org)

### Information management

- Projects,UsersDB Oracle XE APEX [www.oracle.com](www.oracle.com)
- Data Catalogue [http://assda.anu.edu.au/ddiindex.html](http://assda.anu.edu.au/ddiindex.html)

### The client side

-  Any standard web-browser
-  The Kepler Project [www.kepler-project.org](www.kepler-project.org)
-  pgAdmin [www.pgadmin.org](www.pgadmin.org)
-  Git Version Control and GitHub [www.github.com](www.github.com)

# Results

## Case study 1: Historical exposure-response functions

For this case study we replicated the work we had previously conducted
on our personal desktop computers within the University Research
School.  A key result is shown in Figure \ref{fig:Figure1.png}.  That
work was published already (Hanigan \emph{et al.} 2012), however
the improved IT infrastructure offered by this project allows the
analysis to be re-run from a secure web-browser interface.  Such
improved access allowed a much broader discussion of the data,
techniques and results because the researcher was able to discuss the
details of the modelling with other scientists at conferences and
workshops, while actually repeating the computations in real time.
This is a vast improvement over the previous option of leaving the
data analysis on the secure desktop computers at the University, and
merely describing the computations to colleagues at the workshops.

\begin{figure}[!h]
\centering
\includegraphics[width=.45\textwidth]{Figure1.png}
\caption{Drought exposure-response functions Rural Males}
\label{fig:Figure1.png}
\end{figure}

## Case study 2: Future drought scenarios and attributable fraction of suicides

Following the methods of `r citet(bib[["Climate2008"]])` we used the climate change scenarios provided for the Garnaut Review to project estimates of future suicides under various drought conditions.  The statistical method for this calculation is:

$$Y_{ijk}=\sum_{lm}(e^{(\beta_{ijk}} \times { X_{lm})} - 1) \times {BaselineRate_{jkl}} \times { Population_{jklm}}$$

\noindent Where:

$\beta_{ijk}$ = the ExposureVariable coefficient for zone$_i$, age$_j$ and sex$_{k}$

${X_{lm}}$ = Projected Future ExposureVariables

\emph{BaselineRate$_{jkl}$} = \emph{avgDeathsPerTime}/\emph{avgPopPerTime} in age$_j$, sex$_k$ and zone$_l$

\emph{Population$_{jklm}$} = projected populations by age$_j$, sex$_k$, zone$_l$ and time$_m$

# Discussion 
## Principal findings

The results of our project are applicable more generally than just
Reproducible Research (RR). For instance in relation to issues of
governance and management at the multidisciplinary institutions and
multi-institutional projects it is vitally important that data sharing
is enabled, while some data are still being restricted (based either
on authorisation requirements or merely for identifying the profile of
users downloading data for re-use).  It is important that data access
can be made restrictive WITHOUT impeding the progress of the local
science agenda (collaborations, workshops, papers etc) and keeping the
relevant data custodian parties informed about what is happening with
the release of their datasets.  The openness of the analytical
software also has a positive effect on the value of the data
infrastructure (through education and outreach activities) without
risking any unethical or negligent use of these datasets.

#  Discussion and conclusion

- Drought is related to increased suicide risk in Australia
- Future Drought associated deaths can be calculated
- These estimates will be very uncertain, contentious and difficult to justify
- Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration

This system:

- Enables data analysis in a safe environment
- Allows comparison of multiple climate scenarios and assumptions
- Demonstrated with a Climate/Health Impact Assessment 
- This is Reproducible

# References

```{r, echo=FALSE, message=FALSE, eval = T}
write.bibtex(file="references.bib")
```

### 2015-11-04 report1_high_level/manuscript.Rmd
add Garnaut climate change scenarios to the PDF, add the PDF

@@
 ```{r, eval = F, echo = F}
+setwd("~/projects/opensoftware-restricteddata.github.com/report1_high_level/")
 library(rmarkdown)
@@
 render("manuscript.Rmd", "pdf_document")
-#browseURL("manuscript.pdf")
+browseURL("manuscript.pdf")
 ```
@@
 
+```{r, eval = F, echo = F}
+dir()
+dir("~/data")
+indir <- "~/data/GARNAUT mental health archive/1 Data/Rain"
+dir(indir, pattern = "jpg")
+filelist  <- c("A1BDRY_RainSD07.jpg",             "A1BWET_RainSD07.jpg"            )
+for(fi in filelist){
+  #file.copy(file.path(indir, fi), fi)
+  txt <- sprintf("\\begin{figure}[!h]
+\\centering
+\\includegraphics[width=\\textwidth]{%s}
+\\caption{%s}
+\\label{fig:%s}
+\\end{figure}
+", fi, fi, fi)
+  cat(txt)
+}
+```
+
+## The Garnaut review climate change scenarios
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth]{A1BDRY_RainSD07.jpg}
+\caption{A1BDRY RainSD07}
+\label{fig:A1BDRYRainSD07}
+\end{figure}
+
+\begin{figure}[!h]
+\centering
+\includegraphics[width=\textwidth]{A1BWET_RainSD07.jpg}
+\caption{A1BWET RainSD07}
+\label{fig:A1BWET-RainSD07}
+\end{figure}
+
 # Discussion 

### 2015-11-05 report1_high_level/code/droughtIndex_future.R
add drought future func

@@ -1,0 +1,107 @@
+
+# now to setup the drought algorithm to use the thresholds from the
+# historical record, to benchmark the projected future distribution
+
+#### name:droughtIndex_future ####
+droughtIndex<-function(data,years,droughtThreshold=.375){
+# a drought index based on integrated six-monthly rainfall percentiles.
+# based on Professor Mike Hutchinson's work described in
+# Smith D, Hutchinson M, McArthur R. Climatic and Agricultural Drought: Payments and Policy.
+# Canberra, ACT: Centre for Resource and Environmental Studies, Australian National University. 1992.
+
+# Ivan C Hanigan
+# June 2011.
+# GPL2
+# for updates please see https://github.com/ivanhanigan/HutchinsonDroughtIndex.
+
+# my input data are always a data.frame with 4 columns 'date','year','month','rain'
+
+#calculate M month totals
+# started with 6 (current and prior months)
+x<-ts(data[,4],start=1,end=c(years,12),frequency=12)
+x<-c(rep(NA,5),x+lag(x,1)+lag(x,2)+lag(x,3)+lag(x,4)+lag(x,5))
+data$sixmnthtot<-x
+data<-na.omit(data)
+
+# rank in percentage terms with respect to the rainfall totals
+# for the same sequence of 6-months over all years of record
+dataout_final=matrix(nrow=0,ncol=7)
+
+for(i in 1:12){
+x<-data[data$month==i,5]
+#x<-na.omit(x)
+y<-(rank(x)-1)/(length(x)-1)
+# checkpct<-cbind(data[data$month==i,],y)
+# plot(checkpct$sixmnthtot,checkpct$y)
+# rescale between -4 and +4 to replicate palmer index
+z<-8*(y-.5)
+# defualts set the threshold at -1 which is upper limit of mild drought in palmer index (3/8ths, or the 37.5th percentile)
+drought<-x<=quantile(x,droughtThreshold)
+# calculate the drought index for any months that fall below the threshold
+zd<-z*drought
+# save out to the data
+dataout<-data[data$month==i,]
+dataout$index<-z
+dataout$indexBelowThreshold<-zd
+dataout_final=rbind(dataout_final,dataout)
+}
+
+data<-dataout_final[order(dataout_final$date),]
+
+# now calculate the indices
[…]

### 2015-11-05 report1_high_level/code/droughtIndex_future-test.R
add to org

@@ -1,3 +1,4 @@
+
 setwd("report1_high_level/")
 dat <- read.csv("data/rain_future_estimated_dry", stringsAsFactors = F)
 

### 2015-11-05 report1_high_level/code/droughtIndex_future-test.R
Modify droughtIndex function to estimate future drought, based on past distributions

@@ -1,0 +1,35 @@
+setwd("report1_high_level/")
+dat <- read.csv("data/rain_future_estimated_dry", stringsAsFactors = F)
+
+# drop the first year as only half
+names(dat)
+
+head(dat)
+dat$date <- as.Date(paste(dat$year_future, dat$month, 1, sep = "-"))
+
+sds <- names(table(dat$sd_group))
+sds
+par(mfrow=c(2,6))
+sd_drt_out <- matrix(nrow=0,ncol=14)
+for(sd_i in sds){
+#  sd_i <- sds[1]
+dat2 <- dat[dat$year > 1890 & dat$sd_group == sd_i, c('date','year_future','month','avrain','rain_projected')]
+summary(dat2)
+plot(dat2$date, dat2$avrain, type = "l", col='grey') 
+lines(dat2$date, dat2$rain_projected, col = 'blue')
+title(sd_i)
+nyear <- length(names(table(dat2$year_future)))
+nyear
+
+sd_drt <- droughtIndex_future(
+  data=dat2
+  ,
+  years=nyear
+  ,
+  droughtThreshold=.375
+  )
+sd_drt <- data.frame(sd_group = sd_i, sd_drt)
+sd_drt_out <- rbind(sd_drt_out, sd_drt)
+}
+summary(sd_drt_out)
+write.csv(sd_drt_out, "data/drought_future_estimated_dry.csv", row.names = F)

### 2015-11-05 report1_high_level/code/droughtIndex_future.R

@@ -3,7 +3,7 @@
 # historical record, to benchmark the projected future distribution
 
 #### name:droughtIndex_future ####
-droughtIndex<-function(data,years,droughtThreshold=.375){
+droughtIndex_future <- function(data,years,droughtThreshold=.375){
 # a drought index based on integrated six-monthly rainfall percentiles.
 # based on Professor Mike Hutchinson's work described in
 # Smith D, Hutchinson M, McArthur R. Climatic and Agricultural Drought: Payments and Policy.
@@ -14,31 +14,48 @@ droughtIndex<-function(data,years,droughtThreshold=.375){
 # GPL2
 # for updates please see https://github.com/ivanhanigan/HutchinsonDroughtIndex.
 
-# my input data are always a data.frame with 4 columns 'date','year','month','rain'
+# my input data are always a data.frame with 5 columns 'date (future','year(future)','month','rain(past)' 'rain_projected'
 
+#### PAST DISTRIBUTION  
 #calculate M month totals
 # started with 6 (current and prior months)
+  
+# ASSUMES PAST RAIN IS IN FOURTH COLUMN  
 x<-ts(data[,4],start=1,end=c(years,12),frequency=12)
 x<-c(rep(NA,5),x+lag(x,1)+lag(x,2)+lag(x,3)+lag(x,4)+lag(x,5))
 data$sixmnthtot<-x
+#data<-na.omit(data)
+#### FUTURE RAIN
+# ASSUMES FUTURE IS IN COL 5
+x2<-ts(data[,5],start=1,end=c(years,12),frequency=12)
+x2<-c(rep(NA,5),x2+lag(x2,1)+lag(x2,2)+lag(x2,3)+lag(x2,4)+lag(x2,5))
+data$sixmnthtot2<-x2
 data<-na.omit(data)
-
+#head(data)
+#tail(data)
 # rank in percentage terms with respect to the rainfall totals
 # for the same sequence of 6-months over all years of record
 dataout_final=matrix(nrow=0,ncol=7)
 
 for(i in 1:12){
-x<-data[data$month==i,5]
+  #i =1
+  # col sixmnthto is the past rain, sixmnthtot2 is the future rain
+x<-data[data$month==i,'sixmnthtot']
+x2<-data[data$month==i,'sixmnthtot2']
+
 #x<-na.omit(x)
-y<-(rank(x)-1)/(length(x)-1)
[…]

### 2015-11-06 report1_high_level/code/clean_set_datatypes.R
start splitting workflow into modules

@@ -1,0 +1,80 @@
+
+#### name:fit_baseline_model ####
+# create a drought variable for each category
+# ie pre-calculated Drought by Age, Sex and Rural/Urban Region terms, constructed to have the value of the drought index in the specified groups (with Ages grouped by 20 year age brackets) and zero otherwise.
+# NOTE that we initially fitted this model with a drought effect in each 10 year age bracket, however the 20 year age brackets give essentially the same results, and is simpler to calculate.
+  
+require(mgcv)
+require(splines)
+
+# Log transform drought variable, see data preparation for that diagnostic
+data$logDroughtCount = log1p(data$avcount)
+
+# set up the formats of these variables
+data$time=as.Date(paste(data$dthyy,data$dthmm,1,sep='-'))
+data$dthmm=as.factor(data$dthmm)
+data$mm=as.numeric(data$dthmm)
+
+# set up timevar for sinusoidal want
+timevar <- as.data.frame(names(table(data$time)))
+index <- 1:length(names(table(data$time)))
+timevar$time2 <- index/ (length(index) / (length(index)/12))
+names(timevar) <- c('time','timevar')
+timevar$time <- as.Date(timevar$time)
+data <- merge(data,timevar)
+data$time <- as.numeric(data$time)
+data$agegp <- as.factor(data$agegp)
+data$sd_group <- as.factor(data$sd_group)
+str(data)
+
+
+data$rural <-ifelse(data$sd_group %in% c('Central West','Mid-North Coast','Murray','Murrumbidgee','North and Far Western','Northern','Richmond-Tweed','South Eastern'), 1, 0)
+  
+data$agegp2 <-ifelse(data$agegp %in% c('10_19','20_29'), '10_29',
+ifelse(data$agegp %in% c('30_39','40_49'), '30_49',
+ifelse(data$agegp %in% c('50_59','60_69','70plus'), '50plus',
+0)))
+  
+data$agegp2 <- as.factor(data$agegp2)
+  
+ages <- c('10_19','20_29','30_39','40_49','50_59','60_69','70plus')
+ages2 <- c('10_29','30_49','50plus')
+  
+# step thru each
+## for(sexs in 1:2){
+## # sexs <- c(2)#,2)
[…]

### 2015-11-10 report1_high_level/manuscript.Rmd

@@
 
+We can demonstrate the use of this system by using the climate change scenarios held on the database to project out future droughts, and use the confidential suicide data that is safely stored there to estimate baseline risks and future burden of suicide attributable to the droughts.
+Because the server system is easy to access and modify, and these results are reproducible, alternate scenarios and assumptions can be tested.
+
+In the table below, the two rainfall scenarios used by Berry et al 2008 and Bambrick et al 2008 are used to demonstrate this drought impact assessment.
+This shows that the estimated impact of climate change can vary a lot given the input datasets.
+The codes used to fit models, project scenarios and estimate burden of deaths is all available on the system and can be assessed and modified.
+
+```{r, eval = T, echo = F, results = 'asis'}
+library(xtable)
+tab <- read.csv(textConnection("Scenario, Deaths per annum, LCI, UCI
+Historical (1970-2008), 4.01, 2.14, 6.05
+A1FIR1 (Dry), 8.91, 4.56, 14.00
+A1FIR2 (Wet), 2.93, 1.5, 4.47
+"), strip.white = T)
+#tab
+print(xtable(tab), comment = F, include.rownames = F)
+
+```
+
+\clearpage
+
+Below are pictorial represenations of the climate change scenarios influence on rainfall across the country.
+
 \begin{figure}[!h]
 \centering
-\includegraphics[width=\textwidth]{A1BDRY_RainSD07.jpg}
+\includegraphics[width=.5\textwidth]{A1BDRY_RainSD07.jpg}
 \caption{A1BDRY RainSD07}
@@
 \centering
-\includegraphics[width=\textwidth]{A1BWET_RainSD07.jpg}
+\includegraphics[width=.5\textwidth]{A1BWET_RainSD07.jpg}
 \caption{A1BWET RainSD07}

### 2015-12-02 report1_high_level/manuscript.Rmd
add Rmd to the report1 index.org
add stuff for report
modify the PDF link

@@
 partnership', June 2013, Sydney.
-The poster and slideshow are both available to download from this website: [http://opensoftware-restricteddata.github.io/presentations-nccarf-2013/](http://opensoftware-restricteddata.github.io/presentations-nccarf-2013/)
+The poster and slideshow are both available to download from this website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/)
 
@@
 
-\paragraph*{Results:} We built a safe Server/Client Computational
+\paragraph*{Results:} We built a secure Server/Client Computational
 Environment for using open software with restricted data.  We

### 2015-12-02 lib/setup_oracle_of_delphe.r
rm old files from master

@@ -1,468 +1,0 @@
-#################################################################
-# /home/ivan/projects/opensoftware-restricteddata.github.com/lib/setup_oracle_of_delphe.r
-# author:
-# ihanigan
-# date:
-# 2013-08-14
-# description:
-# build tables of oracle of delphe
-#################################################################
-
-# nceph server 03 data management\projects\0.3 Catalogue\backups\f102.sql
-
-# changes
-# 2012-09-17    modify to work on centos linux
-# 2011-06-17    changes made to enable RJDBC on Radha-HP, need to remove ; from sendupdate
-
-# 2010-11-26    add approvals col to accessors for managing ABS mortality URF additional users updates.
-#                               note that ASSDA includes /stdyDscr/dataAccs/
-# 2010-12-01    add table linked to stdyDscr via IDNO called othrStdyMat with relStudy and relPubl
-
-# notes
-# see N:\NCEPH_IT\Data Management\archive\Dataset Library Management\DDILite.xls for mapping decisions re metadata fields
-
-# SSH into a centos box with oracle on it.
-# on shell
-# . /u01/app/oracle/product/11.2.0/xe/bin/oracle_env.sh
-# export LD_LIBRARY_PATH=/u01/app/oracle/product/11.2.0/xe/lib:$LD_LIBRARY_PATH
-## if(!require(RJDBC)) install.packages('RJDBC'); require(RJDBC)
-## drv <- JDBC("oracle.jdbc.driver.OracleDriver",
-##             '/u01/app/oracle/product/11.2.0/xe/jdbc/lib/ojdbc6.jar')
-## ch <- dbConnect(drv,"jdbc:oracle:thin:@115.146.95.169:1521","DDIINDEXDB","passwrd")
-## d <- dbReadTable(ch, name='STDYDSCR', schema = 'DDIINDEXDB')
-## t(d[1,])
-require(devtools)
[…]

### 2015-12-02 report1_high_level/code/clean_set_datatypes.R

@@ -1,80 +1,0 @@
-
-#### name:fit_baseline_model ####
-# create a drought variable for each category
-# ie pre-calculated Drought by Age, Sex and Rural/Urban Region terms, constructed to have the value of the drought index in the specified groups (with Ages grouped by 20 year age brackets) and zero otherwise.
-# NOTE that we initially fitted this model with a drought effect in each 10 year age bracket, however the 20 year age brackets give essentially the same results, and is simpler to calculate.
-  
-require(mgcv)
-require(splines)
-
-# Log transform drought variable, see data preparation for that diagnostic
-data$logDroughtCount = log1p(data$avcount)
-
-# set up the formats of these variables
-data$time=as.Date(paste(data$dthyy,data$dthmm,1,sep='-'))
-data$dthmm=as.factor(data$dthmm)
-data$mm=as.numeric(data$dthmm)
-
-# set up timevar for sinusoidal want
-timevar <- as.data.frame(names(table(data$time)))
-index <- 1:length(names(table(data$time)))
-timevar$time2 <- index/ (length(index) / (length(index)/12))
-names(timevar) <- c('time','timevar')
-timevar$time <- as.Date(timevar$time)
-data <- merge(data,timevar)
-data$time <- as.numeric(data$time)
-data$agegp <- as.factor(data$agegp)
-data$sd_group <- as.factor(data$sd_group)
-str(data)
-
-
-data$rural <-ifelse(data$sd_group %in% c('Central West','Mid-North Coast','Murray','Murrumbidgee','North and Far Western','Northern','Richmond-Tweed','South Eastern'), 1, 0)
-  
-data$agegp2 <-ifelse(data$agegp %in% c('10_19','20_29'), '10_29',
-ifelse(data$agegp %in% c('30_39','40_49'), '30_49',
-ifelse(data$agegp %in% c('50_59','60_69','70plus'), '50plus',
-0)))
-  
-data$agegp2 <- as.factor(data$agegp2)
-  
-ages <- c('10_19','20_29','30_39','40_49','50_59','60_69','70plus')
-ages2 <- c('10_29','30_49','50plus')
-  
-# step thru each
-## for(sexs in 1:2){
-## # sexs <- c(2)#,2)
[…]

### 2015-12-02 report1_high_level/code/do_fit_model.R

@@ -1,33 +1,0 @@
-
-######################
-#do,  The final drought model estimates by age, sex and region
-######################
-# fit the GLM with recommended df
-strt=Sys.time()
-interactionDrtAgeSexRuralModel3 <- glm(deaths ~ sin(timevar*2*pi) + cos(timevar*2*pi)
-+ tmax_anomaly
-+ DrtMales10_29rural
-+ DrtMales30_49rural
-+ DrtMales50plusrural
-+ DrtFemales10_29rural
-+ DrtFemales30_49rural
-+ ns(DrtFemales50plusrural, df = 5)
-+ ns(DrtMales10_29urban, df = 6)
-+ DrtMales30_49urban
-+ ns(DrtMales50plusurban, df = 4)
-+ DrtFemales10_29urban
-+ ns(DrtFemales30_49urban, df = 3)
-+ DrtFemales50plusurban
-+ agegp2
-+ rural
-+ sd_group
-+ sex
-+ agegp
-+ agegp*sex*ns(time,3)
-+ offset(log(pop)), data=data,family=poisson)
-#save.image()
-endd=Sys.time()
-print(endd-strt)
-
-summary(interactionDrtAgeSexRuralModel3)
-#Rsquared.glm.gsm(interactionDrtAgeSexRuralModel3)

### 2015-12-02 report1_high_level/code/droughtIndex_future-test.R

@@ -1,36 +1,0 @@
-
-setwd("report1_high_level/")
-dat <- read.csv("data/rain_future_estimated_dry", stringsAsFactors = F)
-
-# drop the first year as only half
-names(dat)
-
-head(dat)
-dat$date <- as.Date(paste(dat$year_future, dat$month, 1, sep = "-"))
-
-sds <- names(table(dat$sd_group))
-sds
-par(mfrow=c(2,6))
-sd_drt_out <- matrix(nrow=0,ncol=14)
-for(sd_i in sds){
-#  sd_i <- sds[1]
-dat2 <- dat[dat$year > 1890 & dat$sd_group == sd_i, c('date','year_future','month','avrain','rain_projected')]
-summary(dat2)
-plot(dat2$date, dat2$avrain, type = "l", col='grey') 
-lines(dat2$date, dat2$rain_projected, col = 'blue')
-title(sd_i)
-nyear <- length(names(table(dat2$year_future)))
-nyear
-
-sd_drt <- droughtIndex_future(
-  data=dat2
-  ,
-  years=nyear
-  ,
-  droughtThreshold=.375
-  )
-sd_drt <- data.frame(sd_group = sd_i, sd_drt)
-sd_drt_out <- rbind(sd_drt_out, sd_drt)
-}
-summary(sd_drt_out)
-write.csv(sd_drt_out, "data/drought_future_estimated_dry.csv", row.names = F)

### 2015-12-02 report1_high_level/code/droughtIndex_future.R

@@ -1,125 +1,0 @@
-
-# now to setup the drought algorithm to use the thresholds from the
-# historical record, to benchmark the projected future distribution
-
-#### name:droughtIndex_future ####
-droughtIndex_future <- function(data,years,droughtThreshold=.375){
-# a drought index based on integrated six-monthly rainfall percentiles.
-# based on Professor Mike Hutchinson's work described in
-# Smith D, Hutchinson M, McArthur R. Climatic and Agricultural Drought: Payments and Policy.
-# Canberra, ACT: Centre for Resource and Environmental Studies, Australian National University. 1992.
-
-# Ivan C Hanigan
-# June 2011.
-# GPL2
-# for updates please see https://github.com/ivanhanigan/HutchinsonDroughtIndex.
-
-# my input data are always a data.frame with 5 columns 'date (future','year(future)','month','rain(past)' 'rain_projected'
-
-#### PAST DISTRIBUTION  
-#calculate M month totals
-# started with 6 (current and prior months)
-  
-# ASSUMES PAST RAIN IS IN FOURTH COLUMN  
-x<-ts(data[,4],start=1,end=c(years,12),frequency=12)
-x<-c(rep(NA,5),x+lag(x,1)+lag(x,2)+lag(x,3)+lag(x,4)+lag(x,5))
-data$sixmnthtot<-x
-#data<-na.omit(data)
-#### FUTURE RAIN
-# ASSUMES FUTURE IS IN COL 5
-x2<-ts(data[,5],start=1,end=c(years,12),frequency=12)
-x2<-c(rep(NA,5),x2+lag(x2,1)+lag(x2,2)+lag(x2,3)+lag(x2,4)+lag(x2,5))
-data$sixmnthtot2<-x2
-data<-na.omit(data)
-#head(data)
-#tail(data)
-# rank in percentage terms with respect to the rainfall totals
-# for the same sequence of 6-months over all years of record
-dataout_final=matrix(nrow=0,ncol=7)
-
-for(i in 1:12){
-  #i =1
-  # col sixmnthto is the past rain, sixmnthtot2 is the future rain
-x<-data[data$month==i,'sixmnthtot']
-x2<-data[data$month==i,'sixmnthtot2']
-
-#x<-na.omit(x)
-# get distribution of FUTURE RAIN
-y <- (rank(x2)-1)/(length(x2)-1)
-# checkpct<-cbind(data[data$month==i,],y)
[…]

### 2015-12-04 report1_high_level/manuscript.Rmd
some mods

@@
 Australia, especially to the national mortality database after the
-discovery of an incident in which researchers at the University of
-Queensland School of Population Health, dissatisfied with the time
-it was taking the data custodians to provide data, had hacked into
-information about deaths in Australia `r citep(bib[["OKeefe2007"]])`.  
-The subsequent investigations lead to
-a wide ranging modification to the procedures for approval and
-provision of these data that make the access much more restricted.
+discovery of an incident in which Australian population health
+researcher Dr Stephen Begg, dissatisfied with the time it was taking
+the data custodians to provide data, had hacked into information about
+deaths in Australia `r citep(bib[["OKeefe2007"]])`.  The subsequent
+investigations led to a wide ranging modification to the procedures
+for approval and provision of these data that make the access much
+more restricted.
 
-The Replicability Crisis `r citep(bib[["Peng2011"]])` has emerged, reducing public confidence in statements by scientists.
-The true extent of the problem may turn out to be overstated `r citep(bib[["Jager2014"]])` however the concern should be addressed to avoid the problems a lack of confidence in scientific publications would entail.
-Appropriate access to data and analytic software addresses this issue.
-We investigated available workflow tools for data management and analysis and implemented a range of these products on our server.
-This server has enhanced our capacity for experimentation, reviews, revisions and extensions of work in this field.
-We present the results of this project and report that it has streamlined access to population health and environmental data for analysis.
+The Reproducibility Crisis `r citep(bib[["Peng2011"]])` has emerged,
+reducing public confidence in statements by scientists. The true
+extent of the problem may turn out to be overstated
+`r citep(bib[["Jager2014"]])` however the concern should be addressed
+to avoid the problems a lack of confidence in scientific publications
+would entail. Appropriate access to data and analytic software
+addresses this issue.  We investigated available workflow tools for
+data management and analysis and implemented a range of these products
+on our server.  This server has enhanced our capacity for
+experimentation, reviews, revisions and extensions of work in this
+field.  We present the results of this project and report that it has
+streamlined access to population health and environmental data for
+analysis.
 
-## Motivating case study: Climate/suicide
+## Motivating case study: Climate and suicide
 
@@
 However, many papers explore links between suicide and climate variables other than drought (such as temperature) and there are two major reviews papers available of the literature on climatic influences on suicide `r citep(c(bib[["Dixon2009"]],bib[["Deisenhammer2003b"]]))`.
-However, very few studies have investigated the "dryer than average conditions" that is drought specifically.
+However, very few studies have investigated drought specifically.
 

### 2015-12-07 report1_high_level/manuscript.Rmd
A tight edit of the draft manuscript

@@
 ---
-title: 'Open Software - Restricted Data:  A Suicide/Climate Case Study.'
+title: 'Open Software - Restricted Data: A case study.'
 author:  
@@
 
+Draft: \today
+
 \singlespace
@@
 
-\paragraph*{Results:} We built a secure Server/Client Computational
-Environment for using open software with restricted data.  We
+\paragraph*{Results:} We built a secure server/client computational
+environment for using open software with restricted data.  We
 demonstrate the use of this system using drought and suicide as a case study.  
@@
 climate change scenarios with protected intellectual property need to
-be accessed in a restricted way.  In the context of Reproducible
-Research (RR) methods, models and assumptions need to be made
+be accessed in a restricted way.  In the context of reproducible
+research methods, models and assumptions need to be made
 transparent and available for scientific debate even though the
@@
 discovery of an incident in which Australian population health
-researcher Dr Stephen Begg, dissatisfied with the time it was taking
-the data custodians to provide data, had hacked into information about
+researcher Dr Stephen Begg was reported to have 
+hacked into information about
 deaths in Australia `r citep(bib[["OKeefe2007"]])`.  The subsequent
 investigations led to a wide ranging modification to the procedures
-for approval and provision of these data that make the access much
-more restricted.
+for approval and provision of these data that made the access much
+more restricted. No new research applications were approved to access these data between 2009 and 2013.
 
-The Reproducibility Crisis `r citep(bib[["Peng2011"]])` has emerged,
+At the same time the reproducibility crisis `r citep(bib[["Peng2011"]])` has emerged,
 reducing public confidence in statements by scientists. The true
@@
 `r citep(bib[["Jager2014"]])` however the concern should be addressed
-to avoid the problems a lack of confidence in scientific publications
-would entail. Appropriate access to data and analytic software
+to avoid the problems that a lack of confidence in scientific publications
+would entail, especially in respect to evidence-based policy and expenditure of public money. Appropriate access to data and analytic software
 addresses this issue.  We investigated available workflow tools for
@@
 
-## Motivating case study: Climate and suicide
+## Motivating case study: Drought and suicide
 
-Suicide has been linked to climate in a variety of studies.
-The climate change impact on mental health is a gap in knowledge.
-The motivating case for this project was to analyse the relationship between Drought and Suicide. 
-The historical exposure-response function can be used to estimate future climate change impacts.
+The impact of drought on mental health is plausible, however remains a gap in epidemiological knowledge `r citep(bib[["Stanke2013"]])`.
+There is concern too that this health risk will rise under future climate change `r citep(bib[["Berry2008"]])`.
+As mentioned the number of studies that have examined the relationship
+between suicide and drought is limited to only a handful `r citep(c(bib[["Nicholls2006c"]], bib[["Page2002c"]], bib[["Guiney2012a"]], bib[["Hanigan2012e"]]))`.
+The motivating case for this project was to use the historical exposure-response functions to estimate future climate change impacts.
 
@@
 The topic has frequently been raised by the media, by rural politicians and by mental health support groups `r citep(bib[["AustralianB2006"]])`.
-There have also been media reports in India indicating substantial concerns about drought and rural suicide in that country, including `r citep(bib[["SarathiBiswas2012"]])` from May 2012.
+There have also been media reports in India indicating substantial concerns about drought and rural suicide in that country `r citep(bib[["SarathiBiswas2012"]])`.
 
@@
 Such difficulty may occur in conjunction with other economic stresses, such as rising interest rates, falling commodity prices, or an unfavourable foreign exchange rate.
-
-As mentioned the number of studies that have examined the relationship
-between suicide and drought is limited to only a handful `r citep(c(bib[["Nicholls2006c"]], bib[["Page2002c"]], bib[["Guiney2012a"]], bib[["Hanigan2012e"]]))` A  systematic
-literature review of the Health Effects of Drought found little other
-evidence for the putative causal effect of drought on suicide `r citep(bib[["Stanke2013"]])`.
+`r citet(bib[["Vins2015"]])` provide a systematic  literature review of the mental health effects of drought and explore the putative causal mechanisms.
 
@@
 The approach we took to meet the challenge of analysing restricted suicide and climate change scenario data in a safe environment was to
-build a new hardware and software stack to using Open-Source software.  We based our
+build a new hardware and software stack using open-source software.  We based our
 planning on the realisation that there is a growing need of these
-technologies in the context of Reproducible Research (RR).  This
+technologies in the context of reproducible research.  This
 requires that methods, models and assumptions need to be made
@@
 datasets may require authorisation to access.  This is true not just
-in Health data, but also including the context 
-of data with restrictive Intellectual Property and licence requirements.
-
-### Open software for restricted data
+in health data, but also including the context 
+of data  with restrictive intellectual property and licence requirements (such as climate change scenario models).
 
@@
 the issue and analyse the dimensions of the problem we spent the
-initial phase of the project conceptualising an overall "Rich Picture"
+initial phase of the project conceptualising a rich picture
 of the issue, and focused on identifying risks that the project might
@@
 Our design responds directly to the primary threat of unintentional release of
-sensitive data so we decided to build a safe Server/Client environment
+sensitive data so we decided to build a secure server/client environment
 for analysts to develop their software in an open way, while ensuring
@@
 relation to the provision of the server hardware and we were able to
-take advantage of the Nectar Research Cloud for virtualised services.
+take advantage of the Australian 'Nectar Research Cloud' (\url{http://nectar.org.au}) for virtual machines to build the servers on.
 
@@
 data have increased recently, coupled with arguments that appropriate
-access to analytic software is needed to address the Replicability
-Crisis meant that the scope of this project was very broad.  we also
-explored the ambitions of our analysts to support their publishable
-outputs with open software.  Given that examples of un-replicable work
-has spread even to the results published in top journals such as
-Nature and Science, the scope we decided to set for this project was
+access to analytic software is needed to address the reproducibility
+crisis meant that the scope of this project was very broad.  We also
+explored the ambitions of our stakeholders to support their publishable
+outputs with open software.  Given that examples of un-reproducible work
+has spread even to the results published in top journals `r citep(bib[["Peng2015"]])`,
+the scope we decided to set for this project was
 for very high levels of open-ness for the evidence being presented for
-peer-review along with very high levels of restriction on access to
+peer-review, along with very high levels of restriction on access to
 the data. Luckily however we were able to rule out the need for the
-extreme level of restriction such as getting Defence Science and
+extreme level of restriction such as getting Australian Defence Science and
 Technology Organisation (DSTO) accreditation for the security of these
-servers against malicious hacking attacks.
+servers against malicious hacking attacks.  Our servers just needed to be tested by the standard 'vulnerability exploits' scanner used by the Australian National University IT Department.
 
-We also looked at the workflow system Kepler to assess it's utility
+We also looked at the workflow system Kepler to assess it\'s utility
 for providing access to the data, but found that there were a lot of
@@
 different locations, as well as the time needed to test and get user
-acceptance on the services.  Throughout the project we have had to deal
-with lack of resources, especially in terms of Software
-Engineering skills for the Client-side user interface and Linux
-Systems Administration support for the Server-side backend.
+acceptance on the services.
 
-The results of the IT Infrastructure part of this project are described next.
+## System design
 
-## System requirements
-
-In this case study we utilise Virtual Machines (VMs) in the Cloud. Our
+In this case study we utilise Virtual Machines (VMs) in the cloud. Our
 system requires two VMs so that the storage and processing of data can
@@
 Linux commands and configuration specifications are available online at
-[http://opensoftware-restricteddata.github.com](http://opensoftware-restricteddata.github.io).
+[http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/).
 
@@
 
-## Software selection process
+## Software selections
 We researched a variety of systems and found the following set-up worked best for us.
 
-### Linux cluster
+Linux cluster:
 
@@
 
-### PostGIS database
+Geographical Information Systems (GIS) database server:
 
@@
 
-### Analysis
+Statistical analysis server:
 
@@
 
-### Information management
+Information management:
 
@@
 
-### The client side
+The client side:
 
@@
 -  Git Version Control and GitHub [www.github.com](www.github.com)
+-  Emacs code editor with the starter-kit by Kieran Healy and orgmode [http://kieranhealy.org/resources/emacs-starter-kit/](http://kieranhealy.org/resources/emacs-starter-kit/)
 
-# Results
+# Case study
 
-## Case study 1: Historical exposure-response functions
+## Step 1: Historical exposure-response functions
 
 For this case study we replicated the work we had previously conducted
-on our personal desktop computers within the University Research
-School.  A key result is shown in Figure \ref{fig:Figure1.png}.  That
+on our personal desktop computers within the 
+university.  A key result is shown in Figure \ref{fig:Figure1.png}.  That
 work was published already (Hanigan \emph{et al.} 2012), however
@@
 techniques and results because the researcher was able to discuss the
-details of the modelling with other scientists at conferences and
+details of the modelling with other scientists at meetings, conferences and
 workshops, while actually repeating the computations in real time.
 This is a vast improvement over the previous option of leaving the
-data analysis on the secure desktop computers at the University, and
+data analysis on the secure desktop computers at the university, and
 merely describing the computations to colleagues at the workshops.
@@
 
-## Case study 2: Future drought scenarios and attributable fraction of suicides
+## Step 2: Future drought scenarios and attributable fraction of suicides
 
-Following the methods of `r citet(bib[["Climate2008"]])` we used the climate change scenarios provided for the Garnaut Review to project estimates of future suicides under various drought conditions.
+Following the methods of `r citet(bib[["Climate2008"]])` we used the climate change scenarios provided for the Garnaut Climate Change Review `r citep(bib[["CSIRO2013"]])` to project estimates of future suicides under various drought conditions.
 The statistical method for this calculation is:
@@
 
-${X_{lm}}$ = Projected Future ExposureVariables
+${X_{lm}}$ = Projected Future Exposure Variables
 
@@
 
-## The Garnaut review climate change scenarios
+## Results
 
-We can demonstrate the use of this system by using the climate change scenarios held on the database to project out future droughts, and use the confidential suicide data that is safely stored there to estimate baseline risks and future burden of suicide attributable to the droughts.
-Because the server system is easy to access and modify, and these results are reproducible, alternate scenarios and assumptions can be tested.
+In table \ref{tab:table1} below the two rainfall scenarios used by `r citet(bib[["Berry2008"]])` are used to demonstrate this drought impact assessment report.
+For full details of the data from the Garnaut Climate Change Review please refer to the original papers.
+For the purpose of this discussion only a brief summary is required to appreciate the relevance to this case study.
+The result shown below merely shows that the estimated impact of climate change can vary a lot given the input datasets and the assumptions and methods applied during analysis.
+The codes used to fit models, project scenarios and estimate burden of deaths are all available on the system, and can be assessed by request at the project website: \url{http://swish-climate-impact-assessment.github.com/}
 
-In the table below, the two rainfall scenarios used by Berry et al 2008 and Bambrick et al 2008 are used to demonstrate this drought impact assessment.
-This shows that the estimated impact of climate change can vary a lot given the input datasets.
-The codes used to fit models, project scenarios and estimate burden of deaths is all available on the system and can be assessed and modified.
+\bigskip
 
@@
 library(xtable)
-tab <- read.csv(textConnection("Scenario, Deaths per annum, LCI, UCI
+tab <- read.csv(textConnection("Scenario, Deaths per annum, Lower 95th CI, Upper 95th percent CI
 Historical (1970-2008), 4.01, 2.14, 6.05
@@
 #tab
-print(xtable(tab), comment = F, include.rownames = F)
+names(tab) <- gsub("\\.", " ", names(tab))
+print(xtable(tab, caption = 'Estimated number of rural male suicides attributable to drought between 2000 and 2100', label = 'tab:table1'), comment = F, include.rownames = F)
 
@@
 
-Below are pictorial represenations of the climate change scenarios influence on rainfall across the country.
+Below are pictorial represenations of the climate change scenarios estimated annual total rainfall for each Statistical Division (SD) census region across the country.
+SDs are used because this is the geographical unit at which the suicide rates were analysed.
 
@@
 
+We demonstrate the use of this system by using
+the climate change scenarios held on the database to project future
+droughts, and use the confidential suicide data that is safely stored
+there to estimate baseline risks and future burden of suicide
+attributable to the droughts.  Because the server system is easy to
+access through a web browser these results are easily reproducible and
+alternate scenarios and assumptions can be tested.  Because the webserver is very secure, this work can be done with confidence that the risk of unauthorised data release is minimised.
+
 The results of our project are applicable more generally than just
-Reproducible Research (RR). For instance in relation to issues of
+reproducible research. For instance in relation to issues of
 governance and management at the multidisciplinary institutions and
 multi-institutional projects it is vitally important that data sharing
-is enabled, while some data are still being restricted (based either
-on authorisation requirements or merely for identifying the profile of
-users downloading data for re-use).  It is important that data access
+is enabled, while some data are still restricted because access requires
+approval and authorisation, or because data providers wish to be informed of re-use and identifying the profile of
+users of the data.  It is important that data access
 can be made restrictive WITHOUT impeding the progress of the local
@@
 relevant data custodian parties informed about what is happening with
-the release of their datasets.  The openness of the analytical
+the release of their datasets.  The open-ness of the analytical
 software also has a positive effect on the value of the data
@@
 
-#  Discussion and conclusion
+#  Conclusions
 
-- Drought is related to increased suicide risk in Australia
-- Future Drought associated deaths can be calculated
-- These estimates will be very uncertain, contentious and difficult to justify
-- Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration
+In summary we found that:
 
-This system:
+- Climate data can be attached to suicide and census data in a secure computing environment.
+- Future drought associated deaths can be estimated.
+- These estimates are very uncertain, contentious and difficult to justify.
+- Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration that challenges or verifies published results.
 
-- Enables data analysis in a safe environment
-- Allows comparison of multiple climate scenarios and assumptions
-- Demonstrated with a Climate/Health Impact Assessment 
-- This is Reproducible
+This system described:
+
+- Enables data analysis in a safe environment.
+- Allows comparison of multiple climate scenarios and assumptions.
+- Demonstrated with a climate/health impact assessment.
+- Addresses the reproducibility crisis.
 

### 2015-12-07 report1_high_level/manuscript.Rmd
add link for download

@@
 partnership', June 2013, Sydney.
-The poster and slideshow are both available to download from this website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/)
+The poster and slideshow are both available to download from this website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/presentations-nccarf-2013/).
+The latest version of this working paper is available to download from the website: [http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/report1_high_level/manuscript.pdf](http://swish-climate-impact-assessment.github.io/opensoftware-restricteddata/report1_high_level/manuscript.pdf).
 

### 2016-01-23 report1_high_level/manuscript.Rmd
add different image

@@
 
+\singlespace
+
+Unpublished working paper\\
 Draft: \today
 
-\singlespace
-
 \vspace{2mm}\hrule
-\paragraph*{Background:} This essay was written to accompany the material presented as a
+\paragraph*{Background:} This unpublished working paper was written to accompany the material presented as a
 speedtalk and poster at the National Climate Change Adaptation
@@
 
-\paragraph*{Methods:} The paper reports on a project to build tools and procedures for enhancing open and transparent analysis of restricted datasets.
+\paragraph*{Methods:} We report on a project to build tools and procedures for enhancing open and transparent analysis of restricted datasets.
 Some datasets such as suicide or climate
@@
 
-```{r, eval = F, echo = F}
-dir()
-dir("~/data")
-indir <- "~/data/GARNAUT mental health archive/1 Data/Rain"
-dir(indir, pattern = "jpg")
-filelist  <- c("A1BDRY_RainSD07.jpg",             "A1BWET_RainSD07.jpg"            )
-for(fi in filelist){
-  #file.copy(file.path(indir, fi), fi)
-  txt <- sprintf("\\begin{figure}[!h]
-\\centering
-\\includegraphics[width=\\textwidth]{%s}
-\\caption{%s}
-\\label{fig:%s}
-\\end{figure}
-", fi, fi, fi)
-  cat(txt)
-}
-```
-
 ## Results
@@
 \clearpage
+Below are pictorial represenations of the climate change scenarios estimated seasonal rainfall for each Statistical Division (SD) census region across the country.
+SDs are used because this is the geographical unit at which the suicide rates were analysed.
 
-Below are pictorial represenations of the climate change scenarios estimated annual total rainfall for each Statistical Division (SD) census region across the country.
-SDs are used because this is the geographical unit at which the suicide rates were analysed.
+```{r, eval = F, echo = F}
+dir()
+dir("~/data")
+indir <- "~/data/GARNAUT mental health archive/1 Data/Rain/seasonal/"
+dir(indir, pattern = "jpg")
+fi  <- "SeasonalA1FIR1andR2_final.jpg"
+cap <- "Seasonal Rainfall under alternate future predicted scenarios (A1FI) with wet or dry assumptions"
+#file.copy(file.path(indir, fi), fi)
+txt <- sprintf("\\begin{figure}[!h]
+\\centering
+\\includegraphics[width=\\textwidth]{%s}
+\\caption{%s}
+\\label{fig:%s}
+\\end{figure}
+", fi, cap, fi)
+  cat(txt)
+
+```
 
@@
 \centering
-\includegraphics[width=.5\textwidth]{A1BDRY_RainSD07.jpg}
-\caption{A1BDRY RainSD07}
-\label{fig:A1BDRYRainSD07}
+\includegraphics[width=.78\textwidth]{SeasonalA1FIR1andR2_final.jpg}
+\caption{Seasonal rainfall under alternate future predicted scenarios (A1FI) with wet or dry assumptions}
+\label{fig:SeasonalA1FIR1andR2_final.jpg}
 \end{figure}
-
-\begin{figure}[!h]
-\centering
-\includegraphics[width=.5\textwidth]{A1BWET_RainSD07.jpg}
-\caption{A1BWET RainSD07}
-\label{fig:A1BWET-RainSD07}
-\end{figure}
+\clearpage
 
@@
 users of the data.  It is important that data access
-can be made restrictive WITHOUT impeding the progress of the local
+can be made restrictive without impeding the progress of the local
 science agenda (collaborations, workshops, papers etc) and keeping the
@@
 
-In summary we found that:
+In summary we found that the historical weather and climate change data can be attached to suicide and census data in a secure computing environment.
+The future drought associated deaths can then be estimated using statistical tools available in an online virtual machine.
+This facilitates collaboration because the data can be accessed remotely from collaborator\'s offices (or even workshops and conference venues) while adhering to requirements for secure data storage and access constraints.
 
-- Climate data can be attached to suicide and census data in a secure computing environment.
-- Future drought associated deaths can be estimated.
-- These estimates are very uncertain, contentious and difficult to justify.
-- Data management and analysis technology such as that presented is needed to enable rigorous and transparent exploration that challenges or verifies published results.
+In the case study presented it should be noted that these estimates are very uncertain, contentious and difficult to justify.
+However, this reinforces the argument that data management and analysis technology such as those presented are needed to enable rigorous and transparent exploration of methods and assumptions.
+In this way a community of like-minded scientists can challenge or verify published results.
 
-This system described:
-
-- Enables data analysis in a safe environment.
-- Allows comparison of multiple climate scenarios and assumptions.
-- Demonstrated with a climate/health impact assessment.
-- Addresses the reproducibility crisis.
+In conclusion the system described enables data analysis in a safe environment;
+allows comparison of multiple climate scenarios and assumptions;
+is demonstrated with a climate/health impact assessment and addresses the 'reproducibility crisis'.
 
